---
title: "use_cases"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{use_cases}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ROMOPAPI)

test_databasesConfig <- HadesExtras_readAndParseYaml(
  pathToYalmFile = system.file("testdata", "config", "onlyCounts_databasesConfig.yml", package = "ROMOPAPI"),
  pathToFinnGenCountsSqlite = helper_FinnGen_getDatabaseFileCounts()
)

cohortTableHandlerConfig <- test_databasesConfig[[1]]$cohortTableHandler

# Create CDMdbHandler
CDMdbHandler <- HadesExtras_createCDMdbHandlerFromList(cohortTableHandlerConfig, loadConnectionChecksLevel = "basicChecks")
```


# Use case 1: What are the descendants of a concept in the database?

```{r}
getCodeCounts(CDMdbHandler, conceptId = 317009)
```

# Use case 2: How each descendant contributes to the descendant record counts of a concept, over time, and by gender and age?




# Use case 3: What non-standard concepts map to a standard concept and its descendants?




# Use case 4: How each source concept contributes to the descendant record counts of a concept, over time, and by gender and age?